---
title: "1st Year Exam"
author: "Isabel Mejia"
format: html
---

I will use read.csv() to import the COVID-19 variant data from the California Health and Human Services open data site
```{r}
variant_data <- read.csv("covid19_variants.csv")
```

In this next chunk, I will load the approrpiate packages to help with graphing of the data
```{r}
library(lubridate)
library(dplyr)
library(ggplot2)
library(scales)
```

Using as.Date to convert the characters representing the date to the class "Date"

```{r}
variant_data$date <- as.Date(variant_data$date)
```

The data in the data table includes the sum of the variants but I do not want that on my plot so I am going to filter those rows out 
```{r}
variants_only <- filter(variant_data, percentage<100)
head(variants_only)
```

The data has unclassified variants as "Other" but I do not want to plot those so I am going to filter those out as well.
```{r}
variants_name <- filter(variants_only, variant_name != "Other")
head(variants_name)
```

```{r}
variants_name$date <- as.Date(variants_name$date)
```

```{r}
variant_plot <- ggplot(variants_name)+
  aes(date, percentage, group=variant_name, color=variant_name )+
  geom_line()
variant_plot <- variant_plot+labs(x="", y="Percentage of sequenced specimens", title="COVID-19 Variants in California", caption="Data Source: <https://www.cdph.ca.gov/>")
variant_plot <- variant_plot+scale_x_date(date_breaks = "1 month", date_labels = "%b\n%Y")
variant_plot <- variant_plot+ theme(axis.text.x = element_text(angle = 45))
variant_plot
```

